---
title: "Data analysis with simulation of spatial populations"
output: 
  github_document: 
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=T, warning=F, message=F}
library(tidyverse)
library(cowplot)
library(knitr)
library(RcppCNPy)
library(scales)
library(ggrepel)
library(MASS)
```

## PCA with true genotypes

```{bash eval=F}
nohup /programs/plink-2.00/plink2 --pca \
--vcf /workdir/lcwgs-simulation/spatial_pop_sim/rep_1/slim/sample_genotypes.vcf \
--out /workdir/lcwgs-simulation/spatial_pop_sim/rep_1/slim/sample_genotypes \
> /workdir/lcwgs-simulation/nohups/spatial_pop_sim_pca_with_vcf.nohup &
```

```{r eval=T, message=F, warning=F}
eigen_vec <- read_tsv("../spatial_pop_sim/rep_1/slim/sample_genotypes.eigenvec") %>%
  transmute(pc1 = PC1, pc2 = PC2, pop = str_c("p", (ceiling(1:1440/160))), ind = rep(1:160, 9))
eigen_vec %>%
  filter(ind<=80) %>%
  ggplot(aes(x=-pc1, y=-pc2, color=pop)) +
  geom_point() +
  theme_cowplot()
```

## PCA with PCAngsd

### Generate PCA tables

```{r eval=T, warning=F, message=F}
i=1
for (coverage in c(0.125,0.25,0.5,1,2,4)){
  for (sample_size in c(5,10,20,40,80)){
    pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_",coverage,"x.txt")) %>%
      str_extract('p[1-9]')
    ## Read covariance matrix
    cov_matrix <- npyLoad(paste0("../spatial_pop_sim/rep_1/angsd/pcagnsd_bam_list_",sample_size,"_",coverage,"x.cov.npy")) %>%
      as.matrix()
    ## Perform eigen decomposition
    e <- eigen(cov_matrix)
    e_value<-e$values
    x_variance<-e_value[1]/sum(e_value)*100
    y_variance<-e_value[2]/sum(e_value)*100
    e_vector <- as.data.frame(e$vectors)[,1:5]
    pca_table <- bind_cols(pop_label=pop_label, e_vector) %>%
      transmute(population=pop_label, PC1=rescale(V1, c(-1, 1)), PC2=rescale(V2, c(-1, 1)), PC3=rescale(V3, c(-1, 1)), PC4=rescale(V4, c(-1, 1)), PC5=rescale(V5, c(-1, 1)), coverage=coverage, sample_size=sample_size)
    ## Perform DAPC
    fit <- lda(population ~ ., data=pca_table[,1:(5+1)], na.action="na.omit", CV=F, output = "Scatterplot")
    plda <- predict(object = fit,
                  newdata = pca_table[,1:(5+1)])
    prop.lda <- fit$svd^2/sum(fit$svd^2)
    dapc_table <- data.frame(population = pca_table[,1], lda = plda$x) %>%
      transmute(population=population, LD1=rescale(lda.LD1,c(-1,1)), LD2=rescale(lda.LD2, c(-1,1)), coverage=coverage, sample_size=sample_size)
    ## Bind PCA tables and DAPC tables for all sample size and coverage combinations
    if (i==1){
      pca_table_final <- pca_table
      dapc_table_final <- dapc_table
    } else {
      pca_table_final <- bind_rows(pca_table_final,pca_table)
      dapc_table_final <- bind_rows(dapc_table_final,dapc_table)
    }
    i=i+1
  }
}
## Get mean PC values per population
pca_table_final_per_pop <- group_by(pca_table_final, population, coverage, sample_size) %>%
  summarise(PC1_mean=mean(PC1), PC2_mean=mean(PC2), PC3_mean=mean(PC3), PC1_sd=sd(PC1), PC2_sd=sd(PC2), PC3_sd=sd(PC3)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(PC1_mean=rescale(PC1_mean, c(-1,1)), PC2_mean=rescale(PC2_mean, c(-1,1)), PC3_mean=rescale(PC3_mean, c(-1,1)))
## Get mean LD values per population
dapc_table_final_per_pop <- group_by(dapc_table_final, population, coverage, sample_size) %>%
  summarise(LD1_mean=mean(LD1), LD2_mean=mean(LD2), LD1_sd=sd(LD1), LD2_sd=sd(LD2)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(LD1_mean=rescale(LD1_mean, c(-1,1)), LD2_mean=rescale(LD2_mean, c(-1,1)))
```

### Plot PCA

#### PC1 vs. PC2

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(pca_table_final,aes(x=PC1, y=PC2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

#### PC1 vs. PC3

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(pca_table_final,aes(x=PC1, y=PC3, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot PCA centroid

#### PC1 vs. PC2

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(pca_table_final_per_pop, aes(x=PC1_mean, y=PC2_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

#### PC1 vs. PC3

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(pca_table_final_per_pop, aes(x=PC1_mean, y=PC3_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot DAPC

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(dapc_table_final,aes(x=LD1, y=LD2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot DAPC centroid

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(dapc_table_final_per_pop, aes(x=LD1_mean, y=LD2_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

## PCA with PCAngsd, e=2

```{r eval=T, warning=F, message=F}
i=1
for (coverage in c(0.125,0.25,0.5,1,2,4)){
  for (sample_size in c(5,10,20,40,80)){
    pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_",coverage,"x.txt")) %>%
      str_extract('p[1-9]')
    ## Read covariance matrix
    cov_matrix <- npyLoad(paste0("../spatial_pop_sim/rep_1/angsd/pcagnsd_bam_list_",sample_size,"_",coverage,"x_e2.cov.npy")) %>%
      as.matrix()
    ## Perform eigen decomposition
    e <- eigen(cov_matrix)
    e_value<-e$values
    x_variance<-e_value[1]/sum(e_value)*100
    y_variance<-e_value[2]/sum(e_value)*100
    e_vector <- as.data.frame(e$vectors)[,1:5]
    pca_table <- bind_cols(pop_label=pop_label, e_vector) %>%
      transmute(population=pop_label, PC1=rescale(V1, c(-1, 1)), PC2=rescale(V2, c(-1, 1)), PC3=rescale(V3, c(-1, 1)), PC4=rescale(V4, c(-1, 1)), PC5=rescale(V5, c(-1, 1)), coverage=coverage, sample_size=sample_size)
    ## Perform DAPC
    fit <- lda(population ~ ., data=pca_table[,1:(5+1)], na.action="na.omit", CV=F, output = "Scatterplot")
    plda <- predict(object = fit,
                  newdata = pca_table[,1:(5+1)])
    prop.lda <- fit$svd^2/sum(fit$svd^2)
    dapc_table <- data.frame(population = pca_table[,1], lda = plda$x) %>%
      transmute(population=population, LD1=rescale(lda.LD1,c(-1,1)), LD2=rescale(lda.LD2, c(-1,1)), coverage=coverage, sample_size=sample_size)
    ## Bind PCA tables and DAPC tables for all sample size and coverage combinations
    if (i==1){
      pca_table_final <- pca_table
      dapc_table_final <- dapc_table
    } else {
      pca_table_final <- bind_rows(pca_table_final,pca_table)
      dapc_table_final <- bind_rows(dapc_table_final,dapc_table)
    }
    i=i+1
  }
}
## Get mean PC values per population
pca_table_final_per_pop <- group_by(pca_table_final, population, coverage, sample_size) %>%
  summarise(PC1_mean=mean(PC1), PC2_mean=mean(PC2), PC3_mean=mean(PC3), PC1_sd=sd(PC1), PC2_sd=sd(PC2), PC3_sd=sd(PC3)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(PC1_mean=rescale(PC1_mean, c(-1,1)), PC2_mean=rescale(PC2_mean, c(-1,1)), PC3_mean=rescale(PC3_mean, c(-1,1)))
## Get mean LD values per population
dapc_table_final_per_pop <- group_by(dapc_table_final, population, coverage, sample_size) %>%
  summarise(LD1_mean=mean(LD1), LD2_mean=mean(LD2), LD1_sd=sd(LD1), LD2_sd=sd(LD2)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(LD1_mean=rescale(LD1_mean, c(-1,1)), LD2_mean=rescale(LD2_mean, c(-1,1)))
```

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(pca_table_final,aes(x=PC1, y=PC2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

## PCA with PCAngsd, e=8

```{r eval=T, warning=F, message=F}
i=1
for (coverage in c(0.125,0.25,0.5,1,2,4)){
  for (sample_size in c(5,10,20,40,80)){
    pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_",coverage,"x.txt")) %>%
      str_extract('p[1-9]')
    ## Read covariance matrix
    cov_matrix <- npyLoad(paste0("../spatial_pop_sim/rep_1/angsd/pcagnsd_bam_list_",sample_size,"_",coverage,"x_e8.cov.npy")) %>%
      as.matrix()
    ## Perform eigen decomposition
    e <- eigen(cov_matrix)
    e_value<-e$values
    x_variance<-e_value[1]/sum(e_value)*100
    y_variance<-e_value[2]/sum(e_value)*100
    e_vector <- as.data.frame(e$vectors)[,1:5]
    pca_table <- bind_cols(pop_label=pop_label, e_vector) %>%
      transmute(population=pop_label, PC1=rescale(V1, c(-1, 1)), PC2=rescale(V2, c(-1, 1)), PC3=rescale(V3, c(-1, 1)), PC4=rescale(V4, c(-1, 1)), PC5=rescale(V5, c(-1, 1)), coverage=coverage, sample_size=sample_size)
    ## Perform DAPC
    fit <- lda(population ~ ., data=pca_table[,1:(5+1)], na.action="na.omit", CV=F, output = "Scatterplot")
    plda <- predict(object = fit,
                  newdata = pca_table[,1:(5+1)])
    prop.lda <- fit$svd^2/sum(fit$svd^2)
    dapc_table <- data.frame(population = pca_table[,1], lda = plda$x) %>%
      transmute(population=population, LD1=rescale(lda.LD1,c(-1,1)), LD2=rescale(lda.LD2, c(-1,1)), coverage=coverage, sample_size=sample_size)
    ## Bind PCA tables and DAPC tables for all sample size and coverage combinations
    if (i==1){
      pca_table_final <- pca_table
      dapc_table_final <- dapc_table
    } else {
      pca_table_final <- bind_rows(pca_table_final,pca_table)
      dapc_table_final <- bind_rows(dapc_table_final,dapc_table)
    }
    i=i+1
  }
}
## Get mean PC values per population
pca_table_final_per_pop <- group_by(pca_table_final, population, coverage, sample_size) %>%
  summarise(PC1_mean=mean(PC1), PC2_mean=mean(PC2), PC3_mean=mean(PC3), PC1_sd=sd(PC1), PC2_sd=sd(PC2), PC3_sd=sd(PC3)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(PC1_mean=rescale(PC1_mean, c(-1,1)), PC2_mean=rescale(PC2_mean, c(-1,1)), PC3_mean=rescale(PC3_mean, c(-1,1)))
## Get mean LD values per population
dapc_table_final_per_pop <- group_by(dapc_table_final, population, coverage, sample_size) %>%
  summarise(LD1_mean=mean(LD1), LD2_mean=mean(LD2), LD1_sd=sd(LD1), LD2_sd=sd(LD2)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(LD1_mean=rescale(LD1_mean, c(-1,1)), LD2_mean=rescale(LD2_mean, c(-1,1)))
```

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(pca_table_final,aes(x=PC1, y=PC2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

## PCA with covMat

```{r eval=T, warning=F, message=F}
i=1
for (coverage in c(0.125,0.25,0.5,1,2,4)){
  for (sample_size in c(5,10,20,40,80)){
    pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_",coverage,"x.txt")) %>%
      str_extract('p[1-9]')
    ## Read covariance matrix
    cov_matrix <- read_tsv(paste0("../spatial_pop_sim/rep_1/angsd/bam_list_",sample_size,"_",coverage,"x.covMat"), col_names = F) %>%
      as.matrix() %>%
      .[,-(sample_size*9+1)]
    cov_matrix[is.na(cov_matrix)]<- median(cov_matrix, na.rm = T)
    ## Perform eigen decomposition
    e <- eigen(cov_matrix)
    e_value<-e$values
    x_variance<-e_value[1]/sum(e_value)*100
    y_variance<-e_value[2]/sum(e_value)*100
    e_vector <- as.data.frame(e$vectors)[,1:5]
    pca_table <- bind_cols(pop_label=pop_label, e_vector) %>%
      transmute(population=pop_label, PC1=rescale(V1, c(-1, 1)), PC2=rescale(V2, c(-1, 1)), PC3=rescale(V3, c(-1, 1)), PC4=rescale(V3, c(-1, 1)), PC5=rescale(V5, c(-1, 1)), coverage=coverage, sample_size=sample_size)
    ## Perform DAPC
    fit <- lda(population ~ ., data=pca_table[,1:(5+1)], na.action="na.omit", CV=F, output = "Scatterplot")
    plda <- predict(object = fit,
                  newdata = pca_table[,1:(5+1)])
    prop.lda <- fit$svd^2/sum(fit$svd^2)
    dapc_table <- data.frame(population = pca_table[,1], lda = plda$x) %>%
      transmute(population=population, LD1=rescale(lda.LD1,c(-1,1)), LD2=rescale(lda.LD2, c(-1,1)), coverage=coverage, sample_size=sample_size)
    ## Bind PCA tables and DAPC tables for all sample size and coverage combinations
    if (i==1){
      pca_table_final <- pca_table
      dapc_table_final <- dapc_table
    } else {
      pca_table_final <- bind_rows(pca_table_final,pca_table)
      dapc_table_final <- bind_rows(dapc_table_final,dapc_table)
    }
    i=i+1
  }
}
## Get mean PC values per population
pca_table_final_per_pop <- group_by(pca_table_final, population, coverage, sample_size) %>%
  summarise(PC1_mean=mean(PC1), PC2_mean=mean(PC2), PC3_mean=mean(PC3), PC1_sd=sd(PC1), PC2_sd=sd(PC2), PC3_sd=sd(PC3)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(PC1_mean=rescale(PC1_mean, c(-1,1)), PC2_mean=rescale(PC2_mean, c(-1,1)), PC3_mean=rescale(PC3_mean, c(-1,1)))
## Get mean LD values per population
dapc_table_final_per_pop <- group_by(dapc_table_final, population, coverage, sample_size) %>%
  summarise(LD1_mean=mean(LD1), LD2_mean=mean(LD2), LD1_sd=sd(LD1), LD2_sd=sd(LD2)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(LD1_mean=rescale(LD1_mean, c(-1,1)), LD2_mean=rescale(LD2_mean, c(-1,1)))
```

Note: there are more NAs in these covariance matrices compared to the covariance matrices generated from PCAngsd. 

### Plot PCA

#### PC1 vs. PC2

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(pca_table_final,aes(x=PC1, y=PC2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

#### PC1 vs. PC3

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(pca_table_final,aes(x=PC1, y=PC3, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot PCA centroid

#### PC1 vs. PC2

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(pca_table_final_per_pop, aes(x=PC1_mean, y=PC2_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

#### PC1 vs. PC3

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(pca_table_final_per_pop, aes(x=PC1_mean, y=PC3_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot DAPC

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(dapc_table_final,aes(x=LD1, y=LD2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot DAPC centroid

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(dapc_table_final_per_pop, aes(x=LD1_mean, y=LD2_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

## PCoA with ibsMat

```{r eval=T, warning=F, message=F}
i=1
for (coverage in c(0.125,0.25,0.5,1,2,4)){
  for (sample_size in c(5,10,20,40,80)){
    pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_",coverage,"x.txt")) %>%
      str_extract('p[1-9]')
    ## Read covariance matrix
    dist_matrix <- read_tsv(paste0("../spatial_pop_sim/rep_1/angsd/bam_list_",sample_size,"_",coverage,"x.ibsMat"), col_names = F) %>%
      as.matrix() %>%
      .[,-(sample_size*9+1)]
    dist_matrix[is.na(dist_matrix)] <- mean(dist_matrix, na.rm = T)
    ## Perform MDS
    mds <- cmdscale(as.dist(dist_matrix), k=5) %>%
      as.data.frame() 
    mds_table <- bind_cols(pop_label=pop_label, mds) %>%
      transmute(population=pop_label, PCo1=rescale(V1, c(-1, 1)), PCo2=rescale(V2, c(-1, 1)), PCo3=rescale(V3, c(-1, 1)), PCo4=rescale(V4, c(-1, 1)), PCo5=rescale(V5, c(-1, 1)), coverage=coverage, sample_size=sample_size)
    eigen_value <- cmdscale(as.dist(dist_matrix), k=5, eig = T)$eig
    var_explained <- round(eigen_value/sum(eigen_value)*100, 2)

    ## Perform DAPC
    fit <- lda(population ~ ., data=mds_table[,1:(5+1)], na.action="na.omit", CV=F, output = "Scatterplot")
    plda <- predict(object = fit,
                  newdata = mds_table[,1:(5+1)])
    prop.lda <- fit$svd^2/sum(fit$svd^2)
    dapc_table <- data.frame(population = mds_table[,1], lda = plda$x) %>%
      transmute(population=population, LD1=rescale(lda.LD1,c(-1,1)), LD2=rescale(lda.LD2, c(-1,1)), coverage=coverage, sample_size=sample_size)
    ## Bind PCoA tables and DAPC tables for all sample size and coverage combinations
    if (i==1){
      mds_table_final <- mds_table
      dapc_table_final <- dapc_table
    } else {
      mds_table_final <- bind_rows(mds_table_final,mds_table)
      dapc_table_final <- bind_rows(dapc_table_final,dapc_table)
    }
    i=i+1
  }
}
## Get mean PCo values per population
mds_table_final_per_pop <- group_by(mds_table_final, population, coverage, sample_size) %>%
  summarise(PCo1_mean=mean(PCo1), PCo2_mean=mean(PCo2), PCo3_mean=mean(PCo3), PCo1_sd=sd(PCo1), PCo2_sd=sd(PCo2), PCo3_sd=sd(PCo3)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(PCo1_mean=rescale(PCo1_mean, c(-1,1)), PCo2_mean=rescale(PCo2_mean, c(-1,1)), PCo3_mean=rescale(PCo3_mean, c(-1,1)))
## Get mean LD values per population
dapc_table_final_per_pop <- group_by(dapc_table_final, population, coverage, sample_size) %>%
  summarise(LD1_mean=mean(LD1), LD2_mean=mean(LD2), LD1_sd=sd(LD1), LD2_sd=sd(LD2)) %>%
  ungroup() %>%
  group_by(coverage, sample_size) %>%
  mutate(LD1_mean=rescale(LD1_mean, c(-1,1)), LD2_mean=rescale(LD2_mean, c(-1,1)))
```

### Plot PCoA

#### PCo1 vs. PCo2

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(mds_table_final,aes(x=PCo1, y=PCo2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

#### PCo1 vs. PCo3

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(mds_table_final,aes(x=PCo1, y=PCo3, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot PCoA centroid

#### PCo1 vs. PCo2

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(mds_table_final_per_pop, aes(x=PCo1_mean, y=PCo2_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

#### PCo1 vs. PCo3

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(mds_table_final_per_pop, aes(x=PCo1_mean, y=PCo3_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot DAPC

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
ggplot(dapc_table_final,aes(x=LD1, y=LD2, color=population)) +
  geom_point() +
  facet_grid(coverage~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

### Plot DAPC centroid

```{r eval=T, warning=F, message=F, fig.height=12, fig.width=15}
set.seed(1)
ggplot(dapc_table_final_per_pop, aes(x=LD1_mean, y=LD2_mean, color=population, label=population)) +
  geom_point() +
  geom_label_repel(size=3) +
  facet_grid(coverage~sample_size, scales="free") +
  ylim(c(-1.2, 1.2)) +
  xlim(c(-1.2, 1.2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

## PCA with uneven coverage

### PCA with PCAngsd

```{r eval=T, warning=F, message=F, fig.height=2.2, fig.width=10}
i=1
for (sample_size in c(5,10,20,40,80)){
  pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_uneven_coverage.txt")) %>%
    str_extract('p[1-9]')
  coverage <- ifelse(1:(sample_size*9) %% 2 == 1,0.125, 4)
  ## Read covariance matrix
  cov_matrix <- npyLoad(paste0("../spatial_pop_sim/rep_1/angsd/pcagnsd_bam_list_",sample_size,"_uneven_coverage.cov.npy")) %>%
    as.matrix()
  ## Perform eigen decomposition
  e <- eigen(cov_matrix)
  e_value<-e$values
  x_variance<-e_value[1]/sum(e_value)*100
  y_variance<-e_value[2]/sum(e_value)*100
  e_vector <- as.data.frame(e$vectors)[,1:5]
  pca_table <- bind_cols(pop_label=pop_label, e_vector) %>%
    transmute(population=pop_label, PC1=rescale(V1, c(-1, 1)), PC2=rescale(V2, c(-1, 1)), PC3=rescale(V3, c(-1, 1)), PC4=rescale(V4, c(-1, 1)), PC5=rescale(V5, c(-1, 1)), coverage=coverage, sample_size=sample_size)
  ## Bind PCA tables and DAPC tables for all sample size and coverage combinations
  if (i==1){
    pca_table_final <- pca_table
  } else {
    pca_table_final <- bind_rows(pca_table_final,pca_table)
  }
  i=i+1
}
ggplot(pca_table_final,aes(x=PC1, y=PC2, color=population, shape=as.character(coverage))) +
  geom_point() +
  facet_grid(.~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
p1 <- pca_table_final %>%
  mutate(method="PCAngsd") %>%
  ggplot(aes(x=PC1, y=PC2, color=as.character(coverage))) +
  geom_point() +
  facet_grid(method~sample_size, scales="free") +
  scale_color_viridis_d(name = "coverage", option = "D", begin = 0.75, end = 0.1) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "top")
p1
```

### PCA with covMat

```{r eval=T, warning=F, message=F, fig.height=2, fig.width=10}
i=1
for (sample_size in c(5,10,20,40,80)){
  pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_uneven_coverage.txt")) %>%
    str_extract('p[1-9]')
  coverage <- ifelse(1:(sample_size*9) %% 2 == 1,0.125, 4)
  ## Read covariance matrix
  cov_matrix <- read_tsv(paste0("../spatial_pop_sim/rep_1/angsd/bam_list_",sample_size,"_uneven_coverage.covMat"), col_names = F) %>%
    as.matrix() %>%
    .[,-(sample_size*9+1)]
  cov_matrix[is.na(cov_matrix)]<- median(cov_matrix, na.rm = T)
  ## Perform eigen decomposition
  e <- eigen(cov_matrix)
  e_value<-e$values
  x_variance<-e_value[1]/sum(e_value)*100
  y_variance<-e_value[2]/sum(e_value)*100
  e_vector <- as.data.frame(e$vectors)[,1:5]
  pca_table <- bind_cols(pop_label=pop_label, e_vector) %>%
    transmute(population=pop_label, PC1=rescale(V1, c(-1, 1)), PC2=rescale(V2, c(-1, 1)), PC3=rescale(V3, c(-1, 1)), PC4=rescale(V3, c(-1, 1)), PC5=rescale(V5, c(-1, 1)), sample_size=sample_size, coverage=coverage)
  ## Bind PCA tables and DAPC tables for all sample size and coverage combinations
  if (i==1){
    pca_table_final <- pca_table
  } else {
    pca_table_final <- bind_rows(pca_table_final,pca_table)
  }
  i=i+1
}
ggplot(pca_table_final,aes(x=PC1, y=PC2, color=population, shape=as.character(coverage))) +
  geom_point() +
  facet_grid(.~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
p2 <- pca_table_final %>%
  mutate(method="ANGSD (-doCov 1)") %>%
  ggplot(aes(x=PC1, y=PC2, color=as.character(coverage))) +
  geom_point() +
  scale_color_viridis_d(name = "coverage", option = "D", begin = 0.75, end = 0.1) +
  facet_grid(method~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        legend.position = "none")
p2
```

### PCoA with ibsMat

```{r eval=T, warning=F, message=F, fig.height=1.8, fig.width=10}
i=1
for (sample_size in c(5,10,20,40,80)){
  pop_label <- read_lines(paste0("../spatial_pop_sim/rep_1/sample_lists/bam_list_",sample_size,"_uneven_coverage.txt")) %>%
    str_extract('p[1-9]')
  coverage <- ifelse(1:(sample_size*9) %% 2 == 1,0.125, 4)
  ## Read covariance matrix
  dist_matrix <- read_tsv(paste0("../spatial_pop_sim/rep_1/angsd/bam_list_",sample_size,"_uneven_coverage.ibsMat"), col_names = F) %>%
    as.matrix() %>%
    .[,-(sample_size*9+1)]
  dist_matrix[is.na(dist_matrix)] <- mean(dist_matrix, na.rm = T)
  ## Perform MDS
  mds <- cmdscale(as.dist(dist_matrix), k=5) %>%
    as.data.frame() 
  mds_table <- bind_cols(pop_label=pop_label, mds) %>%
    transmute(population=pop_label, PCo1=rescale(V1, c(-1, 1)), PCo2=rescale(V2, c(-1, 1)), PCo3=rescale(V3, c(-1, 1)), PCo4=rescale(V4, c(-1, 1)), PCo5=rescale(V5, c(-1, 1)), coverage=coverage, sample_size=sample_size)
  eigen_value <- cmdscale(as.dist(dist_matrix), k=5, eig = T)$eig
  var_explained <- round(eigen_value/sum(eigen_value)*100, 2)
  ## Bind PCoA tables and DAPC tables for all sample size and coverage combinations
  if (i==1){
    mds_table_final <- mds_table
  } else {
    mds_table_final <- bind_rows(mds_table_final,mds_table)
  }
  i=i+1
}
ggplot(mds_table_final,aes(x=PCo1, y=PCo2, color=population)) +
  geom_point() +
  facet_grid(.~sample_size, scales="free") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
p3 <- mds_table_final %>%
  mutate(method="ANGSD (-doIBS 2)") %>%
  ggplot(aes(x=PCo1, y=PCo2, color=as.character(coverage))) +
  geom_point() +
  facet_grid(method~sample_size, scales="free") +
  scale_color_viridis_d(name = "coverage", option = "D", begin = 0.75, end = 0.1) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")
p3
```

### Assemble plots for batch effect paper

```{r, fig.width=10, fig.height=6}
cowplot::plot_grid(p1, p2, p3, ncol = 1, rel_heights = c(1.3, 1, 1))
```